Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}

### 2017-09-30 paper/abstract.tex
Couple of small changes in abstract, intro and keywords
Added flow scheme, motivation. Implemented comments from Italy.
some changes

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. Data-driven models have been intensively studied in the last year to overcome this problem. However, closed-loop control systems based on such models have not received much attention. We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. We show, using a bilinear model, that DPC has comparable performance with MPC. We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building model, and show that DPC curtails the desired power usage with high confidence. Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants. Our results show that with DPC we obtain an energy saving in terms of $kWh$ that goes from $25.4\%$ to $49.2\%$ when compared to a bang-bang controller.

>>>>

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control. We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. We show, using a bilinear model, that DPC has comparable performance with MPC. We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building model, and show that DPC curtails the desired power usage with high confidence. Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants. Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.

### 2017-09-30 paper/casestudyReal.tex
some changes

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/overview.eps}
		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller. DPC optimzation uses Random Forests with affine functions as predictive models. Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
		\captionsetup{justification=centering}
		\label{F:overview}
	\end{center}
\end{figure}

\subsection{Description of the house}\label{SS:descriptionHouse}
The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}. The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system. Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=24pc]{figures/Vista_sud.eps}
		\caption{Off-grid residential house.}
		\captionsetup{justification=centering}
		\label{F:house}
	\end{center}
\end{figure}

Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building. Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, has a complete independence from the utilities.   
The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes thanks to their high thermal performance. The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.

\begin{figure}[h!]
	\begin{center}
		\subfigure[Walls.]{
			\label{F:houseSection1}
			\centering
			\includegraphics[width=20pc]{figures/blocco_parete_rev01.eps}
		}
		\subfigure[Floor and roof.]{
			\label{F:houseSection2}
			\centering
			\includegraphics[width=20pc]{figures/blocco_solaio_rev01.eps}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{Cross-sections of the wood-cement blocks.}
	\captionsetup{justification=centering}
	\label{F:houseSection}
\end{figure}
The heating system of the building, that supplies the thermal energy required in winter season, is a hydronic system consisting of a vegetable biomass boiler, with a manual ON/OFF, a constant-flow pump station, and tubular steel radiators as shown in Figure \ref{F:housePlantScheme}. The standard biomass boiler has an efficiency of $83.5\%$ with $16.5\ kW$ of thermal power transferred to the water, without gas-flame modulation. The heat transfer fluid distribution is realized through a manifold circuit, that supplies the radiators placed in the various rooms of the house. The energy needs for the domestic hot water (DHW) are covered by the same boiler, coupled with a solar thermal plant. It is worth noting that, in this work, the thermal energy needed for the domestic hot water production is neglected.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=27pc]{figures/Schema_di_impianto.eps}
		\caption{Technological plant scheme of the use case.}
		\captionsetup{justification=centering}
		\label{F:housePlantScheme}
	\end{center}
\end{figure}
For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out. Temperature measuring devices, completely self-produced at the “G. Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of $4$ differently oriented rooms of the house as shown in Figure \ref{F:houseFloors} (displayed with red dots $A1$, $A2$, $A3$ and $A4$). The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house. Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$, displayed with orange dot $G1$) was installed downstream of the biomass boiler to measure the produced thermal energy. The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively. All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
\begin{figure}[h!]
	\begin{center}
		\subfigure[Ground floor.]{
			\label{F:houseGroundFloor}
			\centering
			\includegraphics[width=26pc]{figures/Arch_PT_e_sonde.eps}
		}
		\subfigure[Attic.]{
			\label{F:houseAttic}
			\centering
			\hspace{-1.6cm}
			\includegraphics[width=22.8pc]{figures/Arch_P1.eps}
		}
	\end{center}
	\caption{Layout of the house and probes placement. Legend: red circle for ambient temperature; orange circle for heat meter.}
	\captionsetup{justification=centering}
	\label{F:houseFloors}
\end{figure}

@@

\subsection{EnergyPlus model}\label{SS:energyPlusmodel}
In this section we create an EnergyPlus model of the thermal energy consumption of the house that will be used in Section \ref{SS:simulationResults} to compare the quality of the DPC with respect to a classical bang-bang controller.

To assess energy performance of the use case heating system, the EnergyPlus engine together with DesignBuilder modeling environment has been employed. The model has been created based  on L'Aquila weather data, shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}. According to the K\"{o}ppen-Geiger climate classification, Italy is classified in Csa, Cfa and Csb climate zones, with a warm temperate climate \cite{Peel2007}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
		\caption{Weather data of L'Aquila, Italy.}
		\captionsetup{justification=centering}
		\label{F:houseExternalWeather}
	\end{center}
\end{figure}

Because of the complex morphology of the blocks shown in Figure \ref{F:houseSection}, some simplified assumptions were made, in order to create the EnergyPlus virtual model. Considering the thermal properties of the wood-cement blocks that compose the walls shown in Figure \ref{F:houseSection1}, the thermal effects of the ribs were neglected. For the block employed for floor and roof shown in Figure \ref{F:houseSection2}, a less complex equivalent block, with only three layers, was evaluated in order to consider only one equivalent thermal transmittance value. The properties of the blocks used for the simulation model are listed in Table \ref{T:houseProperties}.

\begin{table}[h!]
	\centering
	\begin{tabular}{ccccc}
		\toprule
		\multirow{3}{1.4cm}{\centering Structural member} & \multirow{3}{3.7cm}{\centering Layer 
															description (from inside to outside)}       & Thermal     		     & Total     & Total        \\
		                                                  &                                             & resistance   		     & thickness & U-value      \\
		                                                  &                                             & $[(m^2K)/W]$		     & $[m]$     & $[W/(m^2K)]$ \\

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=20pc]{figures/Render_rev01.eps}
		\caption{Virtual model of the use case.}
		\captionsetup{justification=centering}
		\label{F:houseVirtualModel}
	\end{center}
\end{figure}

In order to calibrate the EnergyPlus model, a comparison between simulated and measured thermal energy consumption was performed. Considering the off-grid characteristic of the house, the installation of a heat meter was necessary for acquiring the actual energy consumption. Moreover, the heat meter installation has allowed a detailed knowledge of the actual heating system scheduling, shown in Figure \ref{F:houseActualScheduling}, faithfully reproduced in the simulated model. The heat meter, that consists of two thermocouples for flow and return thermal fluid temperatures, a turbine flowmeter and a controller, employed Equation \eqref{Eq:energyConsumption} to calculate the real thermal energy consumption $\dot Q$ of the house.
\begin{equation}\label{Eq:energyConsumption}
\dot Q = 0.2777698\times10^{-3}\times\rho\times\Delta V\times c_p\times\Delta T
\end{equation}
In \eqref{Eq:energyConsumption} $\dot Q$ is the thermal energy consumption $[Wh]$, $\rho$ is the water density $[kg/m^3]$, $\Delta V$ is the water volume variation $[m^3]$ detected by the turbine flowmeter, $c_p = 4.186\, [kJ/(kgK)]$ is the water specific heat at constant pressure, $\Delta T$ is the difference between flow and return temperatures of the water $[K]$, $0.2777698∙10−3$ is a dimensionless conversion factor. The considered period goes from March $15^{th}$ $2016$ to April $15^{th}$ $2016$.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature. From midnight to $8am$ the heating system is always switched off.}
		\captionsetup{justification=centering}
		\label{F:houseActualScheduling}
	\end{center}
\end{figure}
Following the hourly calibration proposed by the M\&V guidelines of ASHRAE \cite{USDOE}, also applied in \cite{Mustafarai2014,Raftery2011}, a simulated model is calibrated when the mean bias error (MBE) and the coefficient of variation of the root mean square error [CV(RMSE)] are less than acceptable tolerances, respectively equal to $\pm 10.0\%$ and $30.0\%$. MBE and CV(RMSE) have been calculated by using Equations \eqref{Eq:mbe} and \eqref{Eq:cvrmse}.

@@

The comparison between numerical and experimental data is shown in Figure \ref{F:houseComparisonExperimental} and shows a quite good agreement. Therefore, with a MBE equal to $7.38\%$ and a CV(RMSE) equal to $8.37\%$, the EnergyPlus model of the use case can be considered well calibrated.

\begin{figure}[h!]
	\begin{center}
		\subfigure[Thermal energy.]{
			\label{F:houseComparisonExperimentalEnergy}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_energy_rev01.eps}
		}
		\subfigure[Thermal power.]{
			\label{F:houseComparisonExperimentalPower}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_power.eps}
		}
	\end{center}
	\caption{Comparison between numerical and experimental data.}
	\captionsetup{justification=centering}
	\label{F:houseComparisonExperimental}
\end{figure}
\subsection{Random forest models}
\label{SS:randomForestsModels}
For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\\\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests. In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors. In $S_2$ we have a model for the power consumption of the house ($\tP$). Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house. For this reason, they are computed to give prediction only for time step $k+1$ and not over the whole horizon $N$. Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}. The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms). The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$). All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$. All the models have been trained on the data from March $11^{th}$ $2016$ to April $26^{th}$ $2016$ and validated on the data from May $1^{st}$ $2016$ to May $15^{th}$ $2016$. The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).

\begin{table}[h!]
	\centering
	\begin{tabular}{cccccc}
		\toprule
		Set       & Power   & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
		\midrule
		$S_1$     & $92.44$ & $96.58$     & $96.99$     & $97.21$     & $96.62$\\
		$S_2$     & $92.21$ & $97.38$     & $97.29$     & $96.93$     & $96.81$\\
		\bottomrule
	\end{tabular}
	\caption{Models accuracy for power and temperatures models in $S_1$ and $S_2$ expressed as $\mathrm{1-NRMSE}\,(\%)$.}
	\captionsetup{justification=centering}
	\label{T:S1accuracy}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=26pc]{figures/power_testing.eps}
	\caption{Power consumption model accuracy validation.}
	\captionsetup{justification=centering}
	\label{F:power_testing}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=26pc]{figures/temperature_testing.eps}
	\caption{Temperature model accuracy validation.}
	\captionsetup{justification=centering}
	\label{F:temperature_testing}
\end{center}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
	\subfigure[Comparison of temperature of room $3$ obtained with DPC and bang-bang controller.]{
		\label{F:temperatures_small}
		\centering
		\includegraphics[width=26pc]{figures/Temperatures_small.eps}
	}
	\subfigure[Input schedules obtained from DPC and bang-bang controller.]{
		\label{F:inputs_small}
		\centering
		\includegraphics[width=26pc]{figures/Inputs_small.eps}
	}
	\end{center}
	\caption{Comparison of DPC and bang-bang control performance.}
	\captionsetup{justification=centering}
	\label{F:comparison_small}
\end{figure}

\paragraph{Result 2} In Figure \ref{F:comparison_all_temperature} a comparison of temperature regulation obtained running the DPC with small, medium and large violations is shown. The results show that with the large violation configuration, that gives more importance to the power consumption minimization than to keep temperature within the bounds, temperature is almost always outside the lower bound during the period when the range is tighter. However the maximum violation is still lower than $1\,\degree C$.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Temperatures_all.eps}
	\end{center}
	\caption{Comparison of DPC control performance in terms of comfort with different violations.}
	\label{F:comparison_all_temperature}
\end{figure}
In Table \ref{T:violationErrors}, MBE and CV(RMSE) violation errors are reported to quantify the bounds violation of DPC, in each of the $3$ configurations, and of bang-bang controller.
\begin{table}[h!]
	\centering
%	\scalebox{0.9}{
		\begin{tabular}{cccccc}
			\toprule
			CONTROLLER  & LBV $\mathrm{MBE}$  & LBV $\mathrm{RMSE}$ & UBV $\mathrm{MBE}$ & UBV $\mathrm{RMSE}$ 	\\ 
			\midrule
			$DPC-SV$    & $0.013$             & $0.028$  			      & $0$    				 & $0$     	  	\\
			$DPC-MV$    & $0.165$ 			  & $0.146$     			  & $0$    				 & $0$		  	\\
			$DPC-LV$    & $0.410$  			  & $0.224$     			  & $0$    				 & $0$	      	\\
			$Bang-bang$ & $0.049$ 			  & $0.080$    				  & $0.0063$ 		     & $0.0120$	  	\\
			\bottomrule
		\end{tabular}
%	}
	\caption{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller.}
	\captionsetup{justification=centering}
	\label{T:violationErrors}
\end{table}
We can see that if we allow very small violations, DPC outperforms bang-bang controller in terms of comfort guarantees.

>>>>

\subsection{Description of the house}\label{SS:descriptionHouse}
The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}. The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system. Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/overview.eps}
		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller. DPC optimzation uses Random Forests with affine functions as predictive models. Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
		\captionsetup{justification=centering}
		\label{F:overview}
	\end{center}
\end{figure}
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=24pc]{figures/Vista_sud.eps}
		\caption{Off-grid residential house.}
		\captionsetup{justification=centering}
		\label{F:house}
	\end{center}
\end{figure}
Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building. Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, has a complete independence from the utilities.   
The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes thanks to their high thermal performance. The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.
\begin{figure}[t!]
	\begin{center}
		\subfigure[Walls.]{
			\label{F:houseSection1}
			\centering
			\includegraphics[width=20pc]{figures/blocco_parete_rev01.eps}
		}
		\subfigure[Floor and roof.]{
			\label{F:houseSection2}
			\centering
			\includegraphics[width=20pc]{figures/blocco_solaio_rev01.eps}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{Cross-sections of the wood-cement blocks.}
	\captionsetup{justification=centering}
	\label{F:houseSection}
\end{figure}
The heating system of the building, that supplies the thermal energy required in winter season, is a hydronic system consisting of a vegetable biomass boiler, with a manual ON/OFF, a constant-flow pump station, and tubular steel radiators as shown in Figure \ref{F:housePlantScheme}. The standard biomass boiler has an efficiency of $83.5\%$ with $16.5\ kW$ of thermal power transferred to the water, without gas-flame modulation. The heat transfer fluid distribution is realized through a manifold circuit, that supplies the radiators placed in the various rooms of the house. The energy needs for the domestic hot water (DHW) are covered by the same boiler, coupled with a solar thermal plant. It is worth noting that, in this work, the thermal energy needed for the domestic hot water production is neglected.

For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out. Temperature measuring devices, completely self-produced at the “G. Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of $4$ differently oriented rooms of the house as shown in Figure \ref{F:houseFloors} (displayed with red dots $A1$, $A2$, $A3$ and $A4$). The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house. Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$, displayed with orange dot $G1$) was installed downstream of the biomass boiler to measure the produced thermal energy. The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively. All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.

@@

\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=27pc]{figures/Schema_di_impianto.eps}
		\caption{Technological plant scheme of the use case.}
		\captionsetup{justification=centering}
		\label{F:housePlantScheme}
	\end{center}
\end{figure}
\begin{figure}[t!]
	\begin{center}
		\subfigure[Ground floor.]{
			\label{F:houseGroundFloor}
			\centering
			\includegraphics[width=26pc]{figures/Arch_PT_e_sonde.eps}
		}
		\subfigure[Attic.]{
			\label{F:houseAttic}
			\centering
			\hspace{-1.6cm}
			\includegraphics[width=22.8pc]{figures/Arch_P1.eps}
		}
	\end{center}
	\caption{Layout of the house and probes placement. Legend: red circle for ambient temperature; orange circle for heat meter.}
	\captionsetup{justification=centering}
	\label{F:houseFloors}
\end{figure}
\subsection{EnergyPlus model}\label{SS:energyPlusmodel}
In this section we create an EnergyPlus model of the thermal energy consumption of the house that will be used in Section \ref{SS:simulationResults} to compare the quality of the DPC with respect to a classical bang-bang controller.

To assess energy performance of the use case heating system, the EnergyPlus engine \cite{energyPlus} together with DesignBuilder modeling environment \cite{designBuilder} has been employed. The model has been created based  on L'Aquila weather data, shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}. According to the K\"{o}ppen-Geiger climate classification, Italy is classified in Csa, Cfa and Csb climate zones, with a warm temperate climate \cite{Peel2007}.

Because of the complex morphology of the blocks shown in Figure \ref{F:houseSection}, some simplified assumptions were made, in order to create the EnergyPlus virtual model. Considering the thermal properties of the wood-cement blocks that compose the walls shown in Figure \ref{F:houseSection1}, the thermal effects of the ribs were neglected. For the block employed for floor and roof shown in Figure \ref{F:houseSection2}, a less complex equivalent block, with only three layers, was evaluated in order to consider only one equivalent thermal transmittance value. The properties of the blocks used for the simulation model are listed in Table \ref{T:houseProperties}.    
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
		\caption{Weather data of L'Aquila, Italy.}
		\captionsetup{justification=centering}
		\label{F:houseExternalWeather}
	\end{center}
\end{figure}
\begin{table}[t!]
	\centering
	\begin{tabular}{ccccc}
		\toprule
		\multirow{3}{1.4cm}{\centering Structural member} & \multirow{3}{3.7cm}{\centering Layer 
															description (from inside to outside)}       & Thermal     		     & Total     & Total        \\
		                                                  &                                             & resistance   		     & thickness & U-value      \\
		                                                  &                                             & $[(m^2K)/W]$		     & $[m]$     & $[W/(m^2K)]$ \\

@@

In order to calibrate the EnergyPlus model, a comparison between simulated and measured thermal energy consumption was performed. Considering the off-grid characteristic of the house, the installation of a heat meter was necessary for acquiring the actual energy consumption. Moreover, the heat meter installation has allowed a detailed knowledge of the actual heating system scheduling, shown in Figure \ref{F:houseActualScheduling}, faithfully reproduced in the simulated model. The heat meter, that consists of two thermocouples for flow and return thermal fluid temperatures, a turbine flowmeter and a controller, employed Equation \eqref{Eq:energyConsumption} to calculate the real thermal energy consumption $\dot Q$ of the house.
\begin{equation}\label{Eq:energyConsumption}
\dot Q = 0.2777698\times10^{-3}\times\rho\times\Delta V\times c_p\times\Delta T
\end{equation}
In \eqref{Eq:energyConsumption} $\dot Q$ is the thermal energy consumption $[Wh]$, $\rho$ is the water density $[kg/m^3]$, $\Delta V$ is the water volume variation $[m^3]$ detected by the turbine flowmeter, $c_p = 4.186\, [kJ/(kgK)]$ is the water specific heat at constant pressure, $\Delta T$ is the difference between flow and return temperatures of the water $[K]$, $0.2777698∙10−3$ is a dimensionless conversion factor. The considered period goes from March $15^{th}$ $2016$ to April $15^{th}$ $2016$.
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=20pc]{figures/Render_rev01.eps}
		\caption{Virtual model of the use case.}
		\captionsetup{justification=centering}
		\label{F:houseVirtualModel}
	\end{center}
\end{figure}
Following the hourly calibration proposed by the M\&V guidelines of ASHRAE \cite{USDOE}, also applied in \cite{Mustafarai2014,Raftery2011}, a simulated model is calibrated when the mean bias error (MBE) and the coefficient of variation of the root mean square error [CV(RMSE)] are less than acceptable tolerances, respectively equal to $\pm 10.0\%$ and $30.0\%$. MBE and CV(RMSE) have been calculated by using Equations \eqref{Eq:mbe} and \eqref{Eq:cvrmse}.

@@

The comparison between numerical and experimental data is shown in Figure \ref{F:houseComparisonExperimental} and shows a quite good agreement. Therefore, with a MBE equal to $7.38\%$ and a CV(RMSE) equal to $8.37\%$, the EnergyPlus model of the use case can be considered well calibrated.  
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature. From midnight to $8am$ the heating system is always switched off.}
		\captionsetup{justification=centering}
		\label{F:houseActualScheduling}
	\end{center}
\end{figure}
\begin{figure}[t!]
	\begin{center}
		\subfigure[Thermal energy.]{
			\label{F:houseComparisonExperimentalEnergy}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_energy_rev01.eps}
		}
		\subfigure[Thermal power.]{
			\label{F:houseComparisonExperimentalPower}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_power.eps}
		}
	\end{center}
	\caption{Comparison between numerical and experimental data.}
	\captionsetup{justification=centering}
	\label{F:houseComparisonExperimental}
\end{figure}
\subsection{Random forest models}
\label{SS:randomForestsModels}
For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\\\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests. In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors. In $S_2$ we have a model for the power consumption of the house ($\tP$). Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house. For this reason, they are computed to give prediction only for time step $k+1$ and not over the whole horizon $N$. Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}. The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms). The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$). All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$. All the models have been trained on the data from March $11^{th}$ $2016$ to April $26^{th}$ $2016$ and validated on the data from May $1^{st}$ $2016$ to May $15^{th}$ $2016$. The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).

\begin{table}[t!]
	\centering
	\begin{tabular}{cccccc}
		\toprule
		Set       & Power   & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
		\midrule
		$S_1$     & $92.44$ & $96.58$     & $96.99$     & $97.21$     & $96.62$\\
		$S_2$     & $92.21$ & $97.38$     & $97.29$     & $96.93$     & $96.81$\\
		\bottomrule
	\end{tabular}
	\caption{Models accuracy for power and temperatures models in $S_1$ and $S_2$ expressed as $\mathrm{1-NRMSE}\,(\%)$.}
	\captionsetup{justification=centering}
	\label{T:S1accuracy}
\end{table}

@@

\begin{figure}[t!]
\begin{center}
	\subfigure[Power consumption model accuracy validation.]{
		\label{F:power_testing}
		\centering
		\includegraphics[width=26pc]{figures/power_testing.eps}
	}
	\subfigure[Temperature model accuracy validation.]{
		\label{F:temperature_testing}
		\centering
		\includegraphics[width=26pc]{figures/temperature_testing.eps}
	}	
	\end{center}
\caption{Model accuracy validation.}
\captionsetup{justification=centering}
\label{F:testing}
\end{figure}

@@

\begin{figure}[t!]
	\begin{center}
	\subfigure[Comparison of temperature of room $3$ obtained with DPC and bang-bang controller.]{
		\label{F:temperatures_small}
		\centering
		\includegraphics[width=26pc]{figures/Temperatures_small.eps}
	}
	\subfigure[Input schedules obtained from DPC and bang-bang controller.]{
		\label{F:inputs_small}
		\centering
		\includegraphics[width=26pc]{figures/Inputs_small.eps}
	}
	\end{center}
	\caption{Comparison of DPC and bang-bang control performance.}
	\captionsetup{justification=centering}
	\label{F:comparison_small}
\end{figure}

\paragraph{Result 2} In Figure \ref{F:comparison_all_temperature} a comparison of temperature regulation obtained running the DPC with small, medium and large violations is shown. The results show that with the large violation configuration, that gives more importance to the power consumption minimization than to keep temperature within the bounds, temperature is almost always outside the lower bound during the period when the range is tighter. However the maximum violation is still lower than $1\,\degree C$.
\begin{figure}[t!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Temperatures_all.eps}
	\end{center}
	\caption{Comparison of DPC control performance in terms of comfort with different violations.}
	\label{F:comparison_all_temperature}
\end{figure}
In Table \ref{T:violationErrors}, MBE and CV(RMSE) violation errors are reported to quantify the bounds violation of DPC, in each of the $3$ configurations, and of bang-bang controller. We can see that if we allow very small violations, DPC outperforms bang-bang controller in terms of comfort guarantees.
